### 2015-12-29 Paper.tex
Write initial version of block bootstrap theory paper

\documentclass[12pt]{article}
\usepackage{amsfonts}
\usepackage{amsmath,amsthm,amssymb,graphicx,setspace,url,booktabs,tabularx}
\usepackage[sort,round]{natbib}
\usepackage[margin=1.25in]{geometry}
\usepackage[small]{caption}

\bibliographystyle{abbrvnat}
\newcommand\citepos[2][]{\citeauthor{#2}'s \citeyearpar[#1]{#2}}

\newtheorem{thm}{Theorem}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{claim}[thm]{Claim}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{res}[thm]{Result}
\newtheorem{lema}{Lemma}[section]
\newtheorem{alg}{Algorithm}
\newtheorem{asmp}{Assumption}[section]

\theoremstyle{definition}

\newtheorem{example}{Example}
\newtheorem{defn}{Definition}
\newtheorem{rem}{Remark}

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\plim}{plim}

\DeclareMathOperator{\const}{const}
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\eqd}{\overset{d}{=}}
\DeclareMathOperator{\ind}{1}
\DeclareMathOperator{\pr}{Pr}
\DeclareMathOperator{\sgn}{sgn}
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\vech}{vech}

\renewcommand{\mod}{\operatorname{mod}}

\newcommand{\clt}{\textsc{clt}}
\newcommand{\fclt}{\textsc{fclt}}
\newcommand{\hac}{\textsc{hac}}
\newcommand{\lln}{\textsc{lln}}
\newcommand{\ned}{\textsc{ned}}

\frenchspacing
\onehalfspacing

\begin{document}

\author{Gray Calhoun\thanks{Economics Department, Iowa State
    University, Ames, IA 50011. Telephone: (515) 294-6271.  Email:
    \texttt{gcalhoun@iastate.edu}. Web: \texttt{http://www.econ.iastate.edu/$\sim$gcalhoun}.}}
\title{Block bootstrap consistency\\under weak assumptions}
\date{\today}
\maketitle

\begin{abstract}\noindent
  This paper weakens the size and moment conditions needed for typical
  block bootstrap methods (i.e. the moving blocks, circular blocks,
  and stationary bootstraps) to be valid for the sample mean of
  Near-Epoch-Dependent functions of mixing processes; they are
  consistent under the weakest conditions that ensure the original
  process obeys a Central Limit Theorem \citep[those
    of][\textit{Econometric Theory}]{Jon:97}.  In doing so, this paper
  extends de Jong's method of proof, a blocking argument, to hold with
  random and unequal block lengths.  This paper also proves that
  bootstrapped partial sums satisfy a Functional CLT under the same
  conditions.

  \noindent \textsc{jel} Classification: C12, C15

  \noindent Keywords: Resampling, Time Series, Near Epoch Dependence,
  Functional Central Limit Theorem
\end{abstract}

\newpage\noindent Block bootstraps, e.g. the moving blocks
\citep{Kun:89,LiS:92}, circular block \citep{PoR:92}, and stationary
bootstraps \citep{PoR:94}, have become popular in Economics, partly
because they do not require the researcher to make parametric
assumptions about the data generating process.  They are valid under
general weak dependence and moment conditions.  Some recent papers
(\citealp{GoW:02}; \citealp{GoJ:03}) relax the dependence and moment
conditions of the original papers to fit with those commonly used in
Econometrics based on Near-Epoch-Dependence
(\ned).\footnote{\citet{GoW:02} show that these bootstrap methods can
  be applied to heterogeneous $L_{2+\delta}$-\ned\ processes of size
  $-2(r-1)/(r-2)$ on a strong mixing sequence of size
  $-r(2+\delta)/(r-2)$, where $r > 2$ and $\delta >0$, when the
  original series has uniformly bounded $3r$-moments.  \citet{GoJ:03}
  relax these conditions to $L_{2+\delta}$-\ned\ of size $-1$ and
  $r+\delta$ moments for the original series, and size
  $-(2+\delta)(r+\delta)/(r-2)$ for the underlying mixing series.
  Both papers require that the expected block length grow with $n$ and
  be $o(n^{1/2})$.}\footnote{An array $\{X_{nt}\}$ is an
  $L_{\rho}$-\ned\ process on a mixing array $\{V_{nt}\}$ if
  \begin{equation}
    \| X_{nt} - \E(X_{nt}
    \mid V_{n,t-m},\dots,V_{n,t+m}) \|_{\rho} \leq d_{nt} v_m
  \end{equation} 
  with $v_m \to 0$ as $m \to \infty$ and $\{d_{nt}\}$ an array of
  positive constants.  It is of size $-\gamma$ if $v_m = O(m^{-\gamma
    - \delta})$ for all $\delta>0$.  Dropping the index ``$n$'' gives
  the series definition.} But these conditions are still stronger than
required for a \clt\ to hold; \citet{Jon:97} has established the \clt\
under $L_2$-\ned\ with smaller size and moment
restrictions.\footnote{\citet{Jon:97} proves that the \clt\ holds for
  averages of $L_2$-\ned\ processes of size $-1/2$ on a strong mixing
  series of size $-r/(r-2)$, $r > 2$ and the original series having
  bounded $r$-moments.}  In this paper, I'll show that these block
bootstrap methods consistently estimate the distribution of the sample
mean under \citepos{Jon:97} assumptions, and show that an \fclt\ holds
as well.  I also relax \citepos{GoW:02} and \citepos{GoJ:03}
requirement that the expected block length be $o(n^{1/2})$ to the
original papers' requirement that it be $o(n)$.

The proof will exploit the conditional independence of the blocks in
each bootstrap.  Each bootstrap proceeds by drawing blocks of $M$
consecutive observations from the original time series, and then
pasting these blocks together to create the new bootstrap time series.
The moving blocks bootstrap does exactly that; the circular block
bootstrap ``wraps'' the observations, so that $(X_{n-1}, X_n, X_1,
X_2)$, for example, is a possible block of length four (letting $X_t$
denote the original time series).  The stationary bootstrap wraps the
observations and also draws $M$ at random for each block;
\citet{PoR:94} suggest drawing $M$ from the geometric distribution.
As the name suggests, the series produced by the stationary bootstrap
are strictly stationary, while those produced by the other methods are
not.

Theorem~\ref{res:C} presents our main result, asymptotic normality of
the distribution of bootstrapped partial sums.  This paper adopt the
standard notation that $\E^{*}$, $\var^{*}$, etc. are the usual
operators with respect to the probability measure induced by the
bootstrap and will use explicit stochastic array notation in this
paper for precision.  Also note that all results are presented for the
scalar case but generalize immediately to vector-valued random
variables.  All of the proofs are presented in the appendix; I only
present proofs for the stationary bootstrap, since proofs for the
other methods are similar and easier.  All limits are taken as $n \to
\infty$ unless otherwise noted.

\begin{thm}\label{res:C}
  Suppose the following conditions hold.
  \begin{enumerate}
  \item $X_{nt}$ is $L_2$-\ned\ of size $-1/2$ on an array $V_{nt}$
    that is either strong mixing of size $-r/(r-2)$ or uniform mixing
    of size $-r/2(r-1)$, with $r > 2$.  The \ned\ magnitude indices
    are denoted $\{d_{nt}\}$.
  \item $\E X_{nt} = \mu_{nt}$, $\mu_{nt} - \bar \mu_n$ is uniformly
    bounded, and $\sqrt{n} \| \bar{X}_{n} - \bar\mu_n \|_2 \to 1$.
  \item There exists an array of positive real numbers $\{c_{nt}\}$
    such that $\max_t c_{nt}$ is uniformly finite, $(X_{nt} -
    \mu_{nt})/c_{nt}$ is uniformly $L_r$-bounded, and $d_{nt}/c_{nt}$
    is uniformly bounded.
  \item $X_{nt}^{*}$ is generated by the stationary bootstrap with
    geometric block lengths with success probability $p_n$, $p_n = c
    n^{-a}$ and $a,c \in (0,1)$, or by the moving or circular block
    bootstrap with block length $M_n$ such that $M_n \to \infty$ and
    $M_n/n \to 0$.
  \end{enumerate}
  Define $\mathcal{B}$ as Brownian Motion and
  $\mathbb{B}_n^{*}(\gamma) = n^{-1/2} \sum_{t=1}^{\lfloor n \gamma
    \rfloor} (X_{nt}^{*} - \mu_{n}^{*})$.  Then
  $\sqrt{n}(\bar{X}_n^{*} - \mu_n^{*}) \to^d N(0,1)$ and
  $\mathbb{B}_n^{*} \to^d \mathcal{B}$.
\end{thm}

Theorem~\ref{res:C} assumes that the series have been normalized by
dividing by the square root of the (population) second moment of
$\sqrt{n} (\bar{X}_n - \bar{\mu}_n)$.  Sometimes, though, it's useful
to treat the variance of $X_{nt}$ as a random variable (call it
$\eta^2_n$).  In such cases the dependence and moment conditions
listed in Theorem~\ref{res:C} apply to the original array and not the
normalized array $\eta^{-1}_n X_{nt}$ as is clear from inspecting the
proofs in this paper and \citet{Jon:97}.\footnote{See also
  \citepos{HaH:80} Theorem 3.2, Corollary 3.1, and following remarks.}
Deriving bounds for $\eta_n^{-1} X_{nt}$ from primitive conditions on
$X_{nt}$ would be difficult.

Note that \citet{Jon:97} allows a little bit more flexibility in his
conditions on the array $\{c_{nt}\}$ \citep[see also][]{Dav:93};
essentially, he allows there to be a single set of blocks with the
maximal $\{c_{nt}\}$ over each block well-behaved, while this
paper requires that his condition hold for every possible partition of
blocks.  This additional restriction is required because the
stationary bootstrap will select the blocks randomly.

Consistency for the sample mean follows as an immediate corollary of
Theorem~\ref{res:C} and \citet[Theorem~2]{Jon:97}.

\begin{cor}\label{res:E} 
  Define $\bar{\mu}_n = n^{-1} \sum_{t=1}^n \mu_{nt}$.  If the
  conditions of Theorem~\ref{res:C} hold then $\sqrt{n}(\bar{X}_n^{*}
  - \mu_n^{*})$ and $\sqrt{n}(\bar{X}_n - \bar{\mu}_n)$ converge in
  distribution to the same limit.
\end{cor}

Corollary~\ref{res:E} justifies using the stationary bootstrap to
conduct inference about $\bar \mu_n$ even though there is considerable
heterogeneity.  This conclusion is also present in \citepos{GoW:02}
and \citepos{GoJ:03} papers, but those authors introduce and emphasize
additional assumptions that ensure the heterogeneity is asymptotically
irrelevant, either because the deviations of $\mu_{nt}$ from $\bar
\mu_n$ are small or because there are a finite number of breaks that
occur in a neighborhood of the first observation.  Such assumptions
are unnecessary.

But heterogeneity rules out a version of Corollary~\ref{res:E} for the
partial sum.  It is not hard to see that under the
assumptions of Theorem~\ref{res:C},
\begin{equation}
  n^{-1/2} \sum_{t=1}^{\lfloor \gamma n \rfloor} \Big(X_{nt}
  - (\gamma n)^{-1} \sum_{s=1}^{\lfloor \gamma n \rfloor}
  \mu_{nt}\Big) \to^d \mathcal{B}_{\xi},
\end{equation}
where $\mathcal{B}_{\xi}$ is a Gaussian process and $\xi$ captures the
asymptotic variance of the partial sums.  It would be uncommon for a
researcher to be interested in these particular random variables; one
would typically be more interested in $\bar \mu_n$ than $(\gamma
n)^{-1} \sum_{s=1}^{\lfloor \gamma n \rfloor} \mu_{nt}$, but these
bootstrap methods do not allow us to approximate the distribution of
$(\gamma n)^{-1/2} \sum_{t=1}^{\lfloor \gamma n \rfloor} (X_{nt} -
\bar \mu_n)$.  The same discussion applies if $\mu_{nt} = \bar{\mu}_n$
for all $t$ but $\mathcal{B}_{\xi} \neq \mathcal{B}$.  Other methods,
such as the local block bootstrap \citep{PaP:02,DPP:03}, may be able to
capture this heterogeneity, but we do not pursue that possibility
further.

For completeness, I'll present a consistency result under the
additional condition that the heterogeneity vanishes asymptotically
\citep[similar to Assumption 2.2 of][]{GoW:02}.  This result is a
corollary of Theorem~\ref{res:C} and \citet[Theorem 3.1]{JoD:00b}.

\begin{cor}\label{res:F} Let $\mathbb{B}_n(\gamma) = n^{-1/2} \sum_{t=1}^{\lfloor n \gamma \rfloor} (X_{nt} -
  \bar{\mu}_{n})$ and suppose the conditions of Theorem~\ref{res:C}
  hold, $\sup_t |\mu_{nt} - \bar{\mu}_n| \to 0$, and
  \begin{equation}
    n^{-1} \sum_{s,t=1}^{\lfloor \gamma n \rfloor} \E (X_{ns} -
    \mu_{ns}) (X_{nt} - \mu_{nt}) \to \gamma
  \end{equation}
  for all $\gamma \in [0,1]$.  Then $\mathbb{B}_n$ and
  $\mathbb{B}_n^{*}$ converge in distribution to the same limit.
\end{cor}

We can develop some more intuition by looking at the argument behind
Theorem~\ref{res:C}.  For concreteness, let $X_{n1},\dots,X_{nn}$
denote the original array and let $X_{n1}^{*},\dots,X_{nn}^{*}$ a
hypothetical series generated by the stationary bootstrap.
Conditional on the number of blocks, $N_n$, and the block lengths,
$M_{n1},\dots,M_{nN_n}$, we can see that
\begin{equation}
  \sqrt{n} \bar{X}^{*}_n = \frac{1}{\sqrt{N_n}} \sum_{i=1}^{N_n}
  \Bigg\{\sqrt{\frac{N_n}{n}} \sum_{t=K_{n,i-1}+1}^{K_{n,i}} X_{nt}^{*} \Bigg\},\qquad\text{with}\quad K_{n0} = 0,
  \quad K_{ni} = \sum_{j=1}^i M_{nj},
\end{equation} 
is the sum of $N_n$ independent random variables.  So $\sqrt{n}
(\bar{X}^{*}_n - \bar{X}_n)$ should obey a \clt\ under very weak moment
conditions and satisfy a relationship like
\begin{equation}
  \label{eq:10}
  \frac{\bar{X}_n^{*} - \mu_n^{*}}{\var^{*}(\bar{X}_n^{*} \mid
    M_{n1},\dots,M_{n,N_n})} \to^{d^{*}} N(0,1).
\end{equation}

This argument applies directly to the moving blocks and circular
bootstraps, since $N_n$ and $M_{ni}$ are deterministic.  But they are
stochastic for the stationary bootstrap and the randomness of $N_n$,
in particular, complicates a direct argument; conditioning on $N_n$
would require us to work with the distributions of $M_{nj}$ given
$M_{n1} + \cdots + M_{n,N_n} = n$.  This information should not matter
in the limit, though, and so, as a first step, I'll show that we can
replace $N_n$ with its expected value without affecting the asymptotic
distribution induced by the bootstrap.

\begin{lem}\label{res:A}
  Consider the following procedure:
  \begin{enumerate}
  \item Draw $M_{n1},\dots,M_{n,\lfloor n p_n \rfloor}$ independently from
    the geometric distribution with success probability $p_n$, let
    $m_n = \sum_{i=1}^{\lfloor n p_n \rfloor} M_{ni}$, and let
    $\mathcal{M}_n = (M_{n1},\dots,M_{n,\lfloor n p_n \rfloor})$.
  \item Draw $\lfloor n p_n \rfloor$ blocks from $X_{n1},\dots,X_{nn}$
    (with the observations ``wrapped'' as in the stationary
    bootstrap), with $M_{ni}$ the length of the $i$th block.
  \end{enumerate}
  Let $\{X_{nt}^{**}\}$ be a hypothetical array generated by this
  procedure, let $\{X_{nt}^{*}\}$ be an array generated by the
  stationary bootstrap, define $\mu^{*}_n = \mu_n^{**} = \bar X_n$,
  and define $\mathbb{B}_n^{**}(\gamma) = m_n^{-1/2}
  \sum_{t=1}^{\lfloor \gamma m_n \rfloor} (X_{nt}^{**} - \mu_n^{**})$.
  Under the conditions of Theorem~\ref{res:C}, $\mathbb{B}_n^{*}$ and
  $\mathbb{B}_n^{**}$ converge to the same limit.
\end{lem}

It is easy to see that if $X_{n1}^*,\dots,X_{nn}^*$ is a hypothetical
sequence generated by the moving or circular blocks bootstrap with
block length $M_n$ and $X_{n1}^{**},\dots,X_{n,m_n}^{**}$ is a
bootstrap generated as in Lemma~\ref{res:A} but with $\lfloor n/M_n
\rfloor$ blocks of length $M_n$, then the conclusion of
Lemma~\ref{res:A} still holds.

Given Lemma~\ref{res:A} and the preceding discussion, consistency for
the distribution of the sample mean can be expected to hold if
\begin{equation}\label{eq:5}
  \frac{\bar{X}^{**}_n - \mu_n^{**}}{\var^{**}(\bar{X}_n^{**} \mid
    \mathcal{M}_n)} \to^{d^{**}} N(0,1)
\end{equation}
and $\var^{**}(m_{n}^{1/2} \bar X_n^{**} \mid \mathcal
\mathcal{M}_n) \to^p 1$.  Lemma~\ref{res:B} establishes a stronger
result necessary for partial sums.

\begin{lem}\label{res:B}  
  If the conditions of Theorem~\ref{res:C} hold then
  $\var^{**}(m_n^{-1/2} \sum_{t=1}^{\lfloor \gamma m_n \rfloor}
  X_{nt}^{**} \mid \mathcal{M}_n) \to^p \gamma$ for all
  $\gamma \in [0,1]$.
\end{lem}

The crux of Lemma~\ref{res:B} is the recognition that
$\var^{**}(m_n^{1/2} \bar{X}_n^{**} \mid \mathcal{M}_n)$ can be written as an
average over squared blocks of consecutive observations of the
original series;\footnote{Obviously, this argument needs to hold for
  the partial sums as well, but I'll just discuss the overall average
  for a clearer presentation.} i.e.
\begin{equation}
  \var^{**}(m_n^{1/2} \bar{X}^{**}_n \mid \mathcal{M}_n) = n^{-1}
  \sum_{\tau=1}^n \sum_{j=1}^{\lfloor n p_n \rfloor} Y_{n, \tau, j}^2
\end{equation}
where $Y_{n, \tau, j} = n^{-1/2}\sum_{t=I_{n,\tau,j-1}+1}^{I_{n,\tau,j}}
(X_{nt} - \bar{X}_n)$ and $0 = I_{n,\tau,0} < I_{n,\tau,1} < \cdots <
I_{n,\tau,\lfloor n p_n \rfloor} = m_n$ (the specifics are given in the
proof and use different notation).  Standard arguments from
\citet{Jon:97} establish that $\sum_j Y_{n,\tau,j}^2 \to^p 1$ for each
$\tau$.

As a final point, the consistency of the bootstrap variance 
follows as a corollary to Lemmas~\ref{res:A} and~\ref{res:B}.

\begin{cor}\label{res:D}
  If the conditions of Theorem~\ref{res:C} hold then
  $\var^{*}(n^{1/2} \bar{X}_n^{*}) = \var(n^{1/2} \bar{X}_n) + o_p$.
\end{cor}

The key insight of this paper is for the stationary bootstrap (other
block bootstraps arise as a special case).  After conditioning on the
block lengths, proving that the partial sums obey an \fclt\ reduces to
proving consistency of the bootstrap variance (conditional on the
block lengths), and the conditional bootstrap variance can be
expressed in terms of blocks of consecutive observations of the
original series.  Since convergence of these blocks is essential to
the \clt\ in general,\footnote{See, for example, \citet{Mcl:74},
  \citet[Chapter 3]{HaH:80}, and \citet[Chapter 24]{Dav:94}.} I
conjecture that the same approach will hold under most forms of
dependence that allow for a \clt\ or \fclt.  Finally, this result
allows for considerable heterogeneity in the original process when
approximating the distribution of a sample mean but not a partial sum.

\appendix

\section*{Appendix: Proofs and Supporting Results}
\setcounter{section}{1}

\begin{lema}\label{res:a4}
  Let $\alpha_n(x)= 1 + ((x-1) \mod n)$, let $\mathcal{N}_n =
  (M_{n1},\dots,M_{n,N_n})$, and suppose that the conditions of
  Theorem~\ref{res:C} hold.  For any $t_0$ and $k$ such that the index
  of the following summations are well defined,
  \begin{equation}\label{eq:15}
    \E^*\Bigg(\sum_{t=t_0+1}^{t_0+k} X_{nt}^* \;\Bigg|\;
    \mathcal{N}_n\Bigg) = \E^{**}\Bigg(\sum_{t=t_0+1}^{t_0+k}
    X_{nt}^{**} \;\Bigg|\; \mathcal{M}_n\Bigg) = k \bar{X}_n,
  \end{equation}
  and
  \begin{align}
    \var^*\Bigg(\sum_{t=t_0+1}^{t_0+k} X_{nt}^* \;\Bigg|\;
    \mathcal{N}_n\Bigg) &= 
    \var^{**}\Bigg(\sum_{t=t_0+1}^{t_0+k} X_{nt}^{**} \;\Bigg|\;
    \mathcal{M}_n\Bigg) \nonumber
    \\& = 
    n^{-1} \sum_{\tau=1}^n \Bigg[ \Bigg\{\sum_{t=t_0 + 1}^{K_{n,\kappa_{n0}}}
       (X_{n,\alpha_n(\tau+t)} - \bar{X}_n)\Bigg\}^2 \nonumber
    \\ &\quad +  
      \sum_{j=\kappa_{n0}+1}^{\kappa_{n1}-1}
       \Bigg\{\sum_{t=K_{n,j -1}+1}^{K_{nj}}
       (X_{n,\alpha_n(\tau+t)} - \bar{X}_n)\Bigg\}^2 \nonumber
    \\ &\quad +
       \Bigg\{\sum_{t=K_{n,\kappa_{n1} - 1} + 1}^{t_0 + k}
       (X_{n,\alpha_n(\tau + t)} - \bar{X}_n) \Bigg\}^2 \Bigg] \label{eq:16} 
  \end{align}
  where $\kappa_{n0}$ and $\kappa_{n1}$ denote the blocks containing $t_0$
  and $t_0 + k$, so $K_{n,\kappa_{n0} - 1} < t_0 \leq K_{n,\kappa_{n0}}$ and
  $K_{n,\kappa_{n1} - 1} < t_0 + k \leq K_{n,\kappa_{n1}}$
\end{lema}

\begin{proof}[Proof of Lemma~\ref{res:a4}]
  Equation~\eqref{eq:15} is immediate.  For~\eqref{eq:16}, notice that
  \begin{multline}
    \var^*\Bigg(\sum_{t=t_0+1}^{t_0+k} X_{nt}^* \;\Bigg|\;
    \mathcal{N}_n\Bigg) = \var^{*}\Bigg( \sum_{t=t_0+1}^{K_{n,\kappa_{n0}}}
    X_{nt}^* \;\Bigg|\; \mathcal{N}_n \Bigg) 
    \\ 
    + \sum_{j=\kappa_{n0} + 1}^{\kappa_{n1} - 1} \var^{*}\Bigg(
    \sum_{t=K_{n,j-1}+1}^{K_{nj}} X_{nt}^* \;\Bigg|\; \mathcal{N}_n
    \Bigg) + \var^{*}\Bigg( \sum_{t=K_{n,\kappa_{n1} - 1}+1}^{t_0 + k}
    X_{nt}^* \;\Bigg|\; \mathcal{N}_n \Bigg),
  \end{multline}
  since the blocks are independent given $\mathcal{N}_n$.  Now, for
  each block,
  \begin{equation}
    \var^{*}\Bigg(
    \sum_{t=K_{n,j-1}+1}^{K_{nj}} X_{nt}^* \;\Bigg|\; \mathcal{N}_n
    \Bigg) = n^{-1} \sum_{\tau=1}^n 
    \Bigg\{\sum_{t=K_{n,j -1}+1}^{K_{nj}} (X_{n,\alpha_n(\tau+t)} -
    \bar{X}_n)\Bigg\}^2,
  \end{equation}
  with similar formulae for the first and last blocks.
  The same argument holds for $\var^{**}$, completing the proof.
\end{proof}

\begin{lema}\label{res:a1}
  Suppose that $M_{n1},\dots,M_{n,\lfloor n p_n \rfloor}$ are
  i.i.d. geometric random variables with success parameter $p_n = c
  n^{-a}$ with $a, c \in (0,1)$, and that $\ell_n = (p_n \log
  p_n^{-1})^{-1}$.  Then (i) $\max_{i=1,\dots,\lfloor n p_n \rfloor}
  M_{ni} / n \to^p 0$ and (ii) $\max_{i=1,\dots,\lfloor n p_n \rfloor}
  M_{ni} / \ell_n^{1+\epsilon} \to^p 0$ as $n \to \infty$ for any
  positive $\epsilon$.
\end{lema}

\begin{proof}[Proof of Lemma~\ref{res:a1}]
  We know that, for any increasing positive sequence $x_n$ such that
  $x_n p_n \to \infty$,
  \begin{equation}
    \pr[\max_i M_{ni} \leq x_n] =
    (1 - (1 - p_n)^{x_n})^{\lfloor n p_n \rfloor} \to \lim \exp(- n p_n (1 - p_n)^{x_n}).
  \end{equation}
  Moreover, $n p_n (1 - p_n)^{x_n} \to \lim n p_n e^{-x_n p_n}$.  To
  prove (i), let $x_n = n x$ for any positive number $x$.  Then
  \begin{equation}
    \pr[\max_i M_{ni}/n \leq x ] \to \lim \exp(-n p_n e^{-n p_n x}) =
    \exp(0) = 1.
  \end{equation}
  Since $x$ is arbitrary, $\max_i M_{ni} / n \to^p 0$.

  For (ii), let $x_n = \ell_n^{1 + \epsilon} x$ and note that 
  \begin{equation}
    p_n \ell_n^{1+\epsilon} \geq p_n^{-(\epsilon-\delta-\epsilon\delta)} =
    c^{-(\epsilon-\delta-\epsilon\delta)}
    n^{a(\epsilon-\delta-\epsilon\delta)} \equiv b n^{a(\epsilon
      - \delta - \epsilon\delta)}
  \end{equation}
  for any $\delta > 0$ and large enough $n$.  Choose $\delta$ small
  enough that $\epsilon > \delta(1 +\epsilon)$. Then
  \begin{equation}
    n p_n \exp(-\ell_n^{1+\epsilon} p_n) \leq n p_n
    \exp(-b n^{a(\epsilon -
      \delta - \epsilon\delta)}) = c
    v_n^{\frac{1-a}{a(\epsilon-\delta-\epsilon\delta)}}
    \exp(-b v_n) \to 0,
  \end{equation}
  with $v_n = n^{a(\epsilon-\delta-\epsilon\delta)}$.  Consequently,
  \begin{equation}
    \pr[\max_i M_{ni}/\ell_n^{1+\epsilon} \leq x ] \to \exp(0) = 1
  \end{equation}
  as well.
\end{proof}

\begin{lema}\label{res:a3}
  Suppose that the conditions of Lemma~\ref{res:A} hold.  Then the
  family of random variables $\big\{\big(k^{-1/2} \sum_{t=t_0 + 1}^{t_0+k}
  X_{nt} \big)^2\big\}_{t_0,k > 0}$ is uniformly integrable.
\end{lema}

\begin{proof}[Proof of Lemma~\ref{res:a3}]
Follows the same argument as in de Jong (1997) or Davidson's papers.
Need to fill in later.
\end{proof}

\begin{lema}\label{res:a4}
  Let $\alpha_n(x)= 1 + ((x-1) \mod n)$ and suppose that the
  conditions of Lemma~\ref{res:A} hold.  For any $t_0,k > 0$,
  \begin{equation}\label{eq:15}
    \E^*\Bigg(\sum_{t=t_0+1}^{t_0+k} X_{nt}^* \;\Bigg|\;
    \mathcal{M}_n\Bigg) = \E^{**}\Bigg(\sum_{t=t_0+1}^{t_0+k}
    X_{nt}^{**} \;\Bigg|\; \mathcal{M}_n\Bigg) = k \bar{X}_n,
  \end{equation}
  and
  \begin{multline}\label{eq:16} 
    \var^*\Bigg(\sum_{t=t_0+1}^{t_0+k} X_{nt}^* \;\Bigg|\;
    \mathcal{M}_n\Bigg) = 
    \var^{**}\Bigg(\sum_{t=t_0+1}^{t_0+k} X_{nt}^{**} \;\Bigg|\;
    \mathcal{M}_n\Bigg) 
    \\
     = n^{-1} \sum_{\tau=1}^n \Bigg[ \Bigg\{\sum_{t=1}^{K_{\kappa_0}-t_0}
       (X_{n,\alpha_n(\tau+t)} - \bar{X}_n)\Bigg\}^2 +  \sum_{j=\kappa_0+1}^{\kappa_1-1}
       \Bigg\{\sum_{t=K_{j -1}-t_0+1}^{K_{j}-t_0}
       (X_{n,\alpha_n(\tau+t)} - \bar{X}_n)\Bigg\}^2  +
       \Bigg\{\sum_{t=K_{\kappa_1 - 1} - t_0 + 1}^{K_{\kappa_1}}
       (X_{nt} - \bar{X}_n) \Bigg\}\Bigg]
  \end{multline}
  (the indices need to be worked out better).
\end{lema}

\begin{proof}[Proof of Lemma~\ref{res:a4}]
  Equation~\eqref{eq:15} is immediate.  For~\eqref{eq:16}, let
  $\kappa_0$ and $\kappa_1$ be the indices of the blocks containing
  $s_0+1$ and $s_0 + k$, i.e. such that $K_{\kappa_0 - 1} < s_0 + 1
  \leq K_{\kappa_0}$ and $K_{\kappa_1 - 1} < s_0 + k \leq
  K_{\kappa_1}$.  Then
  \begin{multline}
    \var^*\Bigg(\sum_{t=t_0+1}^{t_0+k} X_{nt}^* \;\Bigg|\;
    \mathcal{M}_n\Bigg) = \var^{*}\Bigg( \sum_{t=t_0+1}^{K_{\kappa_0}}
    X_{nt}^* \;\Bigg|\; \mathcal{M}_n \Bigg) 
    \\ 
    + \sum_{j=\kappa_0 + 1}^{\kappa_1 - 1} \var^{*}\Bigg(
    \sum_{t=K_{j-1}+1}^{K_{j}} X_{nt}^* \;\Bigg|\; \mathcal{M}_n
    \Bigg) + \var^{*}\Bigg( \sum_{t=K_{\kappa_1 - 1}+1}^{K_{\kappa_1}}
    X_{nt}^* \;\Bigg|\; \mathcal{M}_n \Bigg),
  \end{multline}
  since the blocks are independent given $\mathcal{M}_n$.  Now, for
  each block,
  \begin{equation}
    \var^{*}\Bigg(
    \sum_{\tau=K_{j-1}+1}^{K_{j}} X_{nt}^* \;\Bigg|\; \mathcal{M}_n
    \Bigg) = n^{-1} \sum_{t=1}^n 
    \Bigg\{\sum_{t=K_{\kappa_j -1}+1}^{K_{\kappa_j}} (X_{n,\alpha_n(\tau+t)} -
    \bar{X}_n)\Bigg\}^2,
  \end{equation}
  with similar formulae for the first and last blocks.
  The same argument holds for $\var^{**}$, completing the proof.
\end{proof}

\begin{lema}\label{res:a5}
  Suppose the conditions of Theorem~\ref{res:C} hold, let $\ell_n = (p
  \log p^{-1})^{-1}$, and define $Z_{nj}^C(\tau, \mathcal{M}_n)$,
  $M_{nj}(\tau)$, and $K_{nj}(\tau)$ as in the proof of
  Lemma~\ref{res:B}. Now let $J_n$ be the number of block lengths
  $M_{nj}(\tau)$ that are greater that $\ell_n$; let $j_{n1},\dots,j_{n,J_n}$
  be their indices, so that $j_{n1}$ is the index of the first such
  block, $j_{n2}$ is index of the second, etc.; and define 
  $W_{ni} = (V_{n,I_{i-1}+1},\dots,V_{n,I_i})$
  with $I_0 = 0$ and, for $i > 0$,
  \begin{equation}
    I_i =
    \begin{cases}
      K_{n,j_i} & j_{n,i+1} - j_{ni} = 1, \\ 
      \lfloor (K_{n,j_{n,i+1}} + K_{n,j_{n,i}}) / 2 \rfloor & \text{otherwise.}
    \end{cases}
  \end{equation}

  The, conditional on $\mathcal{M}_n$, $\{Z_{nj_i}^C(\tau,
  \mathcal{M}_n), \mathcal{F}_{nj}\}$ is an $L_2$-mixingale of size
  $-1/2$ with magnitude indices
  \begin{equation}
    C D M_{n,j_i}(\tau) n^{-1} \max(1,B) \max(M_{n,j_i}(\tau)^{1/2}
    \ell_n^{-1/2-\epsilon}, 1)
  \end{equation}
  for some $\epsilon > 0$, with $D \geq 2 \max_t(d_{nt}/c_{nt})
  \max_t(c_{nt})$ and $B$ a finite constant that depends only on the
  mixingale coefficients of $X_{nt}$.\footnote{See
    \citet[Theorem~1.6]{Mcl:75} for details.}
\end{lema}
\begin{proof}[Proof of Lemma~\ref{res:a5}]
  Similar to \citet[Lemma 5]{Jon:97}, we'll show that
  $Z^c_{nj_i}(\tau, \mathcal{M}_n)^2 - Z^c_{nj_i}(\tau,
  \mathcal{M}_n)^2$ is an $L_2$-mixingale (conditional on
  $\mathcal{M}_n$) by first showing that it is $L_2$-\ned\ (also
  conditional on $\mathcal{M}_n$); the conclusion then follows
  immediately.   By the same argument as in de Jong's
  Equation~(A.32), for $k > 1$
  \begin{multline}
    \E\Big( \Big[Z_{nj_i}^C(\tau, \mathcal{M}_n) - \E(
    Z_{nj_i}^C(\tau, \mathcal{M}_n) \mid \mathcal{M}_n;
    W_{n,i-k},\dots,W_{n,i+k}) \Big]^2 \;\big|\; \mathcal{M}_n
    \Big)^{\frac12}\\ \leq k^{-1/2-\epsilon} \cdot C D
    M_{n,j_i}(\tau)^{3/2} n^{-1} \ell_n^{-1/2-\epsilon},
  \end{multline}
  with $\epsilon > 0$ and (for $k = 0$)
  \begin{equation}
    \E\Big(\Big[ Z_{nj_i}^C(\tau, \mathcal{M}_n) - \E (Z_{nj_i}^C(\tau,
    \mathcal{M}_n) \mid \mathcal{M}_n; W_{n,i}) \Big]^2 \;\big|\; \mathcal{M}_n
    \Big)^{\frac12} \\ \leq C D B M_{n,j_i} n^{-1}
  \end{equation}
  where $B$ is a finite constant that depends only on the mixingale
  coefficients of $X_{nt}$ \citep[see][for details]{Mcl:75}.  Note
  that $\{X_{nt}\}$ is a mixingale because it is \ned\ \citep[see,
  e.g.][Theorem 17.5]{Dav:94}.

  Consequently, $Z_{nj_i}^C(\tau, \mathcal{M}_n) - \E(Z_{nj_i}^C(\tau,
  \mathcal{M}_n) \mid \mathcal{M}_n)$ is $L_2$-\ned\ of size $-1/2$,
  and an $L_2$-mixingale of size $-1/2$ as well with magnitude indices
  \begin{equation}
    C D \max(M_{j_i}(\tau)^{3/2} n^{-1}
    \ell_n^{-1/2-\epsilon}, M_{j_i}(\tau) n^{-1}) \max(1, B).
  \end{equation}
  So, using \citet[Theorem 1.6]{Mcl:75} again, we have the bound
  \begin{multline}
    \E\Bigg(\Bigg[ \sum_{i=1}^{J_n} \big(Z_{nj_i}^c(\tau,
    \mathcal{M}_n)^2 - \E(Z_{nj_i}^c(\tau, \mathcal{M}_n)^2 \mid
    \mathcal{M}_n)\big) \Bigg]^2 \;\Big|\; \mathcal{M}_n\Bigg) \\ \leq
    A^2 C^2 \sum_{i=1}^J \max(M_{j_i}(\tau)^{3/2} n^{-1}
    \ell_n^{-1/2-\epsilon}, M_{j_i}(\tau) n^{-1})^2 \equiv A^2 C^2 r_n,
  \end{multline}
  where $A$ is $D \max(1,B)$ multiplied by a constant that only
  depends on the mixingale coefficients.  To finish the proof, notice
  that $r_n \to^p 0$ if $\max_j M_j/n \to^p 0$ and $\max_j M_j /
  \ell_n^{1+\epsilon} \to^p 0$, which both follow from Lemma~\ref{res:a1}.
\end{proof}

\begin{lema}\label{res:a6}
  Define $M_{nj}(\tau)$ and $K_{nj}(\tau)$ as in the proof of
  Lemma~\ref{res:B}.  Under the conditions of Theorem~\ref{res:C}, 
  \begin{equation}
    n^{-1} \sum_{j=2}^{\lfloor \gamma n p_n \rfloor} \E \Bigg(
    \sum_{t=1}^{K_{n,j-1}(\tau)} \sum_{s=K_{n,j-1}(\tau)+1}^{K_{nj}(\tau)}
    (X_{nt} - \bar{\mu}_n) (X_{ns} - \bar{\mu}_n) \;\Big|\;
    \mathcal{M}_n\Bigg) \to^p 0
  \end{equation}
  uniformly in $\gamma$ and $\tau$.
\end{lema}
\begin{proof}[Proof of Lemma~\ref{res:a6}]
  Define $\ell_n = (p_n \log p_n^{-1})^{-1}$.  Then
  \begin{multline}
    n^{-1} \sum_{j=2}^{\lfloor \gamma n p_n \rfloor} \E \Bigg(
    \sum_{t=1}^{K_{n,j-1}(\tau)} \sum_{s=K_{n,j-1}(\tau)+1}^{K_{nj}(\tau)}
    (X_{nt} - \bar{\mu}_n) (X_{ns} - \bar{\mu}_n) \;\Big|\;
    \mathcal{M}_n\Bigg) = O_p(\ell_n p_n) \\
    + n^{-1} \sum_{j=2}^{\lfloor \gamma n p_n \rfloor} \E \Bigg(
    \sum_{t=1}^{K_{n,j-1}(\tau)}
    \sum_{s=K_{n,j-1}(\tau)+\ell_n+1}^{K_{nj}(\tau)} (X_{nt} - \bar{\mu}_n)
    (X_{ns} - \bar{\mu}_n) \;\Bigg|\; \mathcal{M}_n\Bigg)
  \end{multline}
  uniformly in $\tau$ and $\gamma$ by \citet[Theorem~1.6]{Mcl:75}.
  Then $\ell_n p_n \to 0$ by construction and the remainder of the
  proof follows the same argument as \citepos{Jon:97} Lemma~4.
\end{proof}

\subsection*{Proof of Theorem~\ref{res:C}}
By Lemma~\ref{res:A}, it suffices to prove that $\mathbb{B}_n^{**}$
obeys the \fclt.  We have
\begin{multline}\label{eq:1}
  m_n^{-1/2} \sum_{t=1}^{\lfloor \gamma m_n \rfloor} (X_{nt}^{**} -
  \mu_n^{**}) = (n p)^{-1/2} \sum_{j=1}^{\lfloor \gamma n p_n \rfloor}
  \Big\{(n p_n/m_n)^{1/2} \sum_{t = K_{n,j-1} + 1}^{K_{nj}}
  (X_{nt}^{**} - \mu_n^{**})\Big\} \\ + m_n^{-1/2} \sgn(\gamma m_n -
  K_{n,\lfloor \gamma n p_n \rfloor}) \sum_{t = \min(\lfloor \gamma
    m_n \rfloor, K_{n,\lfloor \gamma n p_n \rfloor}) +
    1}^{\max(\lfloor \gamma m_n \rfloor, K_{n,\lfloor \gamma n p_n
      \rfloor})} (X_{nt}^{**} - \mu_n^{**}).
\end{multline}
We can show that the second sum in~(\ref{eq:1}) vanishes
asymptotically.  To simplify the presentation, assume that $\gamma m_n
> K_{n,\lfloor \gamma n p_n \rfloor}$; the same argument works for the
reverse inequality as well.  By Lemma~\ref{res:a4},
\begin{equation}
  \E^{**} m_n^{-1/2} \sum_{t = K_{n,\lfloor \gamma n p_n \rfloor} +
    1}^{\lfloor \gamma m_n\rfloor} (X_{nt}^{**} - \mu_n^{**})=
  (\bar X_n - \bar X_n) \E^{**} (\lfloor \gamma m_n
  \rfloor - K_{n,\lfloor \gamma n p_n \rfloor})/ m_n^{1/2} = 0
\end{equation}
(remember that $m_n$ is random) and
\begin{align}
  \var^{**}\Big(&m_n^{-1/2} \sum_{t = K_{n,\lfloor \gamma n p_n
      \rfloor} + 1}^{\lfloor \gamma m_n\rfloor} (X_{nt}^{**} -
  \mu_n^{**})\Big) \\&= \E^{**} (m_n n)^{-1} \sum_{\tau=1}^n
  \Bigg\{\sum_{j=\lfloor\gamma n p_n \rfloor + 1}^{\kappa_n - 1}
  \Bigg(\sum_{t=K_{n,j-1}+1}^{K_{nj}} (X_{nt} - \bar{X}_n)\Bigg)^2 \\
  &\qquad+ \Bigg(\sum_{t=K_{n,\kappa_n-1}+1}^{\lfloor \gamma m_n
    \rfloor} (X_{nt} - \bar{X}_n)\Bigg)^2 \Bigg\} \to^p 0
\end{align}
uniformly in $\gamma$ by \citet[Theorem~1.6]{Mcl:75}, with $\kappa_n$
indicating the block containing $\lfloor \gamma m_n \rfloor$, so
$K_{n,\kappa_n-1} < \lfloor \gamma m_n \rfloor \leq K_{n,\kappa_n}$.
It now suffices to prove that the first term in~(\ref{eq:1}) satisfies
an \fclt.

Let $Z_{nj}^{**} = \sqrt{n p_n/ m_n} \sum_{t = K_{n,j-1} + 1}^{K_{nj}}
(X_{nt}^{**} - \mu_n^{**})$; clearly $Z_{nj}^{**}$ is an \mds\
conditional on $\mathcal{M}_n$ that has finite variance and is
globally covariance stationarity condition by Lemma~\ref{res:B}.
Moreover, $(n p_n)^{-1} \sum_j (Z_{nj}^{**})^2 \to^{p^{**}} 1$ by the
\lln\ \citep[e.g.][Theorem~19.7]{Dav:94} and $\max_j (Z_{nj}^{**})^2 =
o_{p^{**}}(n)$ given $\mathcal{M}_n$ by uniform integrability.  Then
the random functions $\gamma \mapsto (n p_n)^{-1/2}
\sum_{j=1}^{\lfloor \gamma n p_n \rfloor} Z_{nj}^{**}$ converge in
distribution to $\mathcal{B}$ conditional on $\mathcal{M}_n$ by the
\fclt\ \citep[e.g.][Theorem 27.14]{Dav:94}.  Since this limit does not
depend on $\mathcal{M}_n$, it holds unconditionally as well.\qed

\subsection*{Proof of Lemma~\ref{res:A}}
We'll use a coupling argument.  If $m_n > n$, then $\bar{X}_n^{*} \eqd
\bar{Y}_n^{**} \equiv n^{-1} \sum_{t=1}^{n} X_{nt}^{**}$.  Similarly,
if $m_n < n$, $\bar{X}_n^{**} \eqd \bar{Z}_n^{*} \equiv n^{-1}
\sum_{t=1}^{m_n} X_{nt}^{*}$.  So the result holds if $m_n^{1/2}
(\bar{X}_n^{**} - \bar{Y}_n^{**}) \to^p 0$ and $n^{1/2} (\bar{X}_n^{*}
- \bar{Z}_n^{*}) \to^p 0$; i.e.
\begin{equation}\label{eq:4}
  m_n^{-1/2} \sum_{t=\lfloor \gamma m_n \rfloor +
    1}^{\lfloor \gamma n \rfloor} (X_{nt}^{*} - \mu_n^{*}) \to^{p}
  0, \qquad (n^{-1/2} - m_n^{-1/2})
  \sum_{t=1}^{\lfloor \gamma n \rfloor} (X_{nt}^* - \mu_n^*) \to^p 0
\end{equation}
and
\begin{equation}\label{eq:6}
  n^{-1/2} \sum_{t=\lfloor \gamma n \rfloor +
    1}^{\lfloor \gamma m_n \rfloor} (X_{nt}^{**} - \mu_n^{**}) \to^{p} 0, \qquad
  (m_n^{-1/2} - n^{-1/2})
  \sum_{t=1}^{\lfloor \gamma m_n \rfloor} (X_{nt}^{**} - \mu_n^{**}) \to^p 0
\end{equation}
uniformly in $\gamma$, where summations over empty index sets are
defined to be zero.  I'll prove~\eqref{eq:4} as the argument
for~\eqref{eq:6} is identical

Now let $\mathcal{N}_n = (M_{n1},\dots,M_{n,N_n})$.  By
Lemma~\ref{res:a4},
\begin{gather}\label{eq:17}
  \E^{*} \Big(m_n^{-1/2} \sum_{t=\lfloor \gamma m_n \rfloor +
    1}^{\lfloor \gamma n \rfloor} (X_{nt}^{*} - \mu_n^{*}) \;\Big|\;
  \mathcal{N}_n \Big) = \frac{(\lfloor \gamma n \rfloor - \lfloor
    \gamma m_n
    \rfloor)^{+}}{n^{1/2}} (\bar{X}_n - \bar{X}_n) = 0, \\
  \E^{*} \Big((n^{-1/2} - m_n^{-1/2}) \sum_{t=1}^{\lfloor \gamma n
    \rfloor} (X_{nt}^* - \mu_n^*) \;\Big|\; \mathcal{N}_n \Big) = 0,
\end{gather}
and
\begin{align}
  \var^{*} \Big(m_n^{-1/2} & \sum_{t=\lfloor \gamma m_n \rfloor +
    1}^{\lfloor \gamma n \rfloor} (X_{nt}^{*} - \mu_n^{*}) \;\Big|\;
  \mathcal{N}_n \Big) \\ &= (m_n n)^{-1} \sum_{\tau=1}^n \Bigg[
  \Bigg\{ \sum_{t= \lfloor \gamma m_n \rfloor + 1}^{K_{n,\kappa_{n0}}}
  (X_{n,\alpha_n(\tau + t)} - \bar{X}_n) \Bigg\}^2 \\&\quad +
  \sum_{j=\kappa_{n0} + 1}^{\kappa_{n1} - 1} \Bigg\{ \sum_{t=K_{n,j-1}
    + 1}^{K_{nj}} (X_{n, \alpha_n(\tau+t)} - \bar X_n)\Bigg\}^2
  \\&\quad + \Bigg\{ \sum_{t=K_{n,\kappa_{n1} - 1} + 1}^{\lfloor
    \gamma n \rfloor} (X_{n, \alpha_n(\tau + t)} - \bar{X}_n)\Bigg\}^2
  \Bigg].
\end{align}
By \citet[Theorem~1.6]{Mcl:75},
\begin{multline}
  \E \var^{*} \Big( (\gamma m_n)^{-1/2} \sum_{t=\lfloor \gamma m_n
    \rfloor + 1}^{\lfloor \gamma n \rfloor} (X_{nt}^* - \mu_n^*)
  \;\Big|\; \mathcal{N}_n\Big) \\ \leq \gamma^{-1} \E (m_{n} n)^{-1} \sum_{\tau=1}^n
  \E \Big[ A_{n0} (K_{n,\kappa_{n0}} - \lfloor \gamma m_n \rfloor) \\
  + \sum_{j=\kappa_{n0} + 1}^{\kappa_{n1} - 1} A_{nj} M_{nj} +
  A_{n,\kappa_{n1}} (\lfloor \gamma n \rfloor - K_{n,\kappa_{n1}} - 1)
  \Big] \to 0
\end{multline}
uniformly in $\gamma$, since the $A_{nj}$ are uniformly bounded
constants that only depend on the mixingale magnitude indices of
$\{X_{nt}\}$.  Also,
\begin{equation}
  \var^{*} \Big((n^{-1/2} - m_n^{-1/2})
  \sum_{t=1}^{\lfloor \gamma n \rfloor} (X_{nt}^* - \mu_n^*) \mid
  \mathcal{N}_n\Big) \to^p 0
\end{equation}
uniformly by a similar argument.  \citet[Theorem~1.6]{Mcl:75} also
ensures that the conditional variances are uniformly integrable,
giving~\eqref{eq:4}.\qed

\subsection*{Proof of Lemma~\ref{res:B}}
For any $\tau$ in $\{1,\dots,n\}$, let $\kappa_n(\tau)$ indicate the
block containing $\tau$, so $K_{n,\kappa_n(\tau)-1} < \tau \leq
K_{n,\kappa_n(\tau)}$.  Also let
\begin{equation}
  M_{nj}(\tau) = (\min(n, K_{n,\beta_n(\kappa_n(\tau) + j)} - \tau) - \max(1,
  K_{n,\beta_n(\kappa_n(\tau) - 1 + j)} + 1 - \tau))^{+},
\end{equation}
with $\beta_n(x) = (x-1) \mod \lfloor n p_n \rfloor + 1$, and let
$K_{nj}(\tau) = \sum_{i=1}^j M_{nj}(\tau)$ and $K_{n0}(\tau) = 0$.
Now define
\begin{equation}
  Z_{nj}(\tau,\mathcal{M}_n) = \sum_{t=K_{n,j-1}(\tau)+1}^{K_{nj}(\tau)} n^{-1/2} (X_{nt} - \bar \mu_n), \qquad j = 1,\dots,
  \lfloor \gamma n p_n \rfloor.
\end{equation}
By Lemma~\ref{res:a4},
\begin{equation}
  \var^{**}\Big(m_n^{-1/2} \sum_{t=1}^{\lfloor \gamma m_n\rfloor} X_{nt}^{**} \mid \mathcal{M}_n\Big) = n^{-1}
  \sum_{\tau = 1}^n \sum_{j=1}^{\lfloor \gamma n p_n \rfloor} Z_{nj}(\tau,
  \mathcal{M}_n)^2 + o_p
\end{equation}
uniformly in $\gamma$.  The advantage of this representation is that
each series $\{Z_{nj}(\tau, \mathcal{M})\}_j$ is formed from
consecutive blocks of the original series.

Define the lower bound $\ell_n$ on the lengths $M_{ni}$ as $\ell_n =
(p_n \log p_n^{-1})^{-1}$.  Since
\begin{multline}
  \sum_{j=1}^{\lfloor \gamma n p_n \rfloor} \E\big( Z_{nj}(\tau,
  \mathcal{M}_n)^2 \ind\{M_{nj}(\tau) > \ell_n\}\big) \leq \\ \big((1
  - p_n)^{1/p_n}\big)^{p_n\ell_n} \sum_{j=1}^{\lfloor \gamma n p_n
    \rfloor} \E Z_{nj}(\tau, \mathcal{M}_n)^2 = e^{-p_n\ell_n} O(1)
  \to 0,
\end{multline}
with the last equality a consequence of \citet[Theorem~1.6]{Mcl:75},
we have
\begin{equation}
  \var^{**}(m_n^{1/2} \bar{X}^{**}_n \mid \mathcal{M}_n) = n^{-1} \sum_{\tau=1}^n
  \sum_{j=1}^{\lfloor \gamma n p_n \rfloor} Z_{nj}(\tau, \mathcal{M}_n)^2
  \ind\{M_{nj}(\tau) > \ell_n\} + o_p.
\end{equation}

Now, truncate $Z_{nj}(\tau, \mathcal{M}_n)$ similar to
\citet[Lemma~5]{Jon:97}: define
\begin{equation}\label{eq:3}
  Z^C_{nj}(\tau, \mathcal{M}_n) = 
  \begin{cases}
    0 & M_{nj}(\tau) \leq \ell_n, \\ Z_{nj}(\tau, \mathcal{M}_n) &
    M_{nj}(\tau) > \ell_n ,\ |Z_{nj}(\tau, \mathcal{M}_n)| \leq C
    \sqrt{M_{nj}(\tau)/n} \\ C \sqrt{M_{nj}(\tau)/n} & M_{nj}(\tau) >
    \ell_n ,\ Z_{nj}(\tau, \mathcal{M}_n) >C \sqrt{M_{nj}(\tau)/n} \\
    -C \sqrt{M_{nj}(\tau)/n} & M_{nj}(\tau) > \ell_n,\ Z_{nj}(\tau,
    \mathcal{M}_n) < -C \sqrt{M_{nj}(\tau)/n}.
  \end{cases}
\end{equation}
\citet[Theorem~1.6]{Mcl:75} establishes that the family $\{n
Z_{nj}(\tau, \mathcal{M})^2 / M_{nj}(\tau); n,j,\tau,\mathcal{M}\}$ is
uniformly integrable,\footnote{See also \citet[Lemma 3.2]{Dav:92} and
  \citet[Lemma 5]{Jon:97}} so $Z_{nj}(\tau,\mathcal{M}_n)^2 -
Z^c_{nj}(\tau,\mathcal{M}_n)^2$ can be made arbitrarily small by
choosing a large enough $C$.  So it suffices to prove
\begin{equation}\label{eq:7}
  \sum_{j=1}^{\lfloor \gamma n p_n \rfloor} \big(Z_{nj}^C(\tau,
  \mathcal{M}_n)^2 - \E(Z_{nj}^C(\tau,
  \mathcal{M}_n)^2 \mid \mathcal{M}_n) \big) \to^p 0
\end{equation}
and
\begin{equation}\label{eq:8}
  n^{-1} \sum_{\tau=1}^n \sum_{j=1}^{\lfloor \gamma n p_n \rfloor}
  \E(Z_{nj}(\tau, \mathcal{M}_n)^2 \mid \mathcal{M}_n) \to^p
  \gamma. 
\end{equation}

\citet[Theorem~1.6]{Mcl:75} and Lemma~\ref{res:a5} imply that, for
some $\epsilon > 0$,
\begin{multline}
  \E\Bigg(\Bigg\{\sum_{j=1}^{\lfloor \gamma n p_n \rfloor}
  \big(Z_{nj}^C(\tau, \mathcal{M}_n)^2 - \E(Z_{nj}^C(\tau,
  \mathcal{M}_n)^2 \mid \mathcal{M}_n) \big)\Bigg\}^2 \;\Big|\;
  \mathcal{M}_n \Bigg) \\ \leq A^2 C^2 \sum_{j=1}^{\lfloor \gamma n
    p_n \rfloor} (M_{nj}(\tau)/n)^2 \max(1, M_{nj}(\tau) /
  \ell_n^{1+\epsilon}),
\end{multline}
which converges to zero if $\max_j M_{nj}/n \to^p 0$ and $\max_j
M_{nj}/\ell_n^{1+\epsilon} \to^p 0$.  Both follow from
Lemma~\ref{res:a1}, proving~\eqref{eq:7}.

To prove~(\ref{eq:8}), let $d_n$ be a sequence of integers such that
$d_n \to \infty$ and $d_n/n \to \infty$ and let $\zeta_n(\gamma)$ be a
sequence of functions such that $\zeta_n(x) = \zeta_n(x - 1)$ for all
$x$ and
\begin{equation}
  (2 d_n)^{-1} \sum_{s,t = \lfloor \gamma n \rfloor - d_n +
    1}^{\lfloor \gamma n \rfloor + d_n} \E (X_{ns} -
  \bar{\mu}_n)(X_{nt} - \bar{\mu}_n) - \zeta_n(\gamma) \to 0
\end{equation}
for all $\gamma \in (0,1]$. It is obvious that \eqref{eq:8} holds for
$\gamma = 0$.  Lemmas~\ref{res:a5} and~\ref{res:a6} and the fact that
$\sum_{\tau=1}^n n^{-1} \zeta_n(x + \tau/n) \to 1$ uniformly in $x$ by
assumption imply that
\begin{align}
  n^{-1} \sum_{\tau=1}^n \sum_{j=1}^{\lfloor \gamma n p_n \rfloor}
  \E(Z_{nj}^2(\tau, \mathcal{M}_n) \mid \mathcal{M}_n) &=
  \sum_{\tau=1}^n \sum_{j=1}^{\lfloor \gamma n p_n \rfloor}
  \frac{\zeta_n(\tau/n + j/np_n)}{n^2 p_n} + o_p \\ &= \gamma + o_p,
\end{align}
completing the proof.  \qed

\bibliography{AllRefs}
\end{document}

% LocalWords:  CLT Kun LiS PoR GoW GoJ nt indices eq studentized JoD reindex nj
% LocalWords:  de Jong's ns Mcl AllRefs nn th nm ni HaH formulae jn gcalhoun nN
% LocalWords:  jel PaP DPP np