### 2015-12-29 subs/ET/reply_editor.txt
Add response letters to referees and editor

Dear Robert,

Thank you again for giving me the opportunity to revise my submission to
Econometric Theory, "Block bootstrap consistency under weak assumptions"
(manuscript #2868).

As you mention in your letter, two of the referees have concerns about the
execution of the proofs in the paper, in particular (in the case of referee
3) technical concerns about uniformity that arise from conditioning on the
sequence of block lengths in the Stationary Bootstrap.  To address these
concerns, I have substantially rewritten the proofs: I had originally used
mixingale arguments to justify the CLT and the FCLT and argued that the
bootstrapped process was a mixingale "conditional on" an additional
information set, but that concept is nonstandard, causing confusion, and
introduces many subtle technical issues that referee 3 alerted me to.  The
revised paper uses instead a martingale difference sequence CLT for the
bootstrapped sample mean and shows that the CLTs assumptions are justified
by deriving the necessary inequalities from NED properties of the original
process.  This change leads to a more direct and more standard logical
argument and I hope addresses the referees' concerns.

The referees also point out that the conclusions of the main results of the
paper could be presented more precisely.  In the revised paper, I present
the CLT and the FCLT results as separate theorems and make my claims about
"convergence in distribution" much more precise: see Theorems 1 and 2 for
the revised statements.  I should point out that referee 1 had the main
concern:

"The above-cited papers, in common with most of this literature, prove
uniform convergence in probability of the difference between the conditional
probability measure of the statistic computed from the resampled data and
the corresponding distribution in the original data.... This is not done
here."

This concern comes from my presentation of the results but not from the
underlying math (in particular, since the limit is continuous, the uniform
convergence that the referee desires is a direct consequence of convergence
in distribution).  I have changed the conclusions to make this uniform
convergence explicit and explain the connection in the proof of Theorem 1.
I should also note that this notion of "uniformity" is distinct from the
technical issues that concern referee #3 (referee #3 is concerned about
uniformity over parameters related to events in the conditioning sigma-field,
not over the domain of the CDF), and the mixingale/MDS changes I mention
above are needed to address referee #3.

Finally, I have made many other changes to the paper's proofs to try to
improve its clarity.  The most substantial of these changes is the removal
of an intermediate approximation: the original version of the paper
approximated the bootstrap process $X_t^*$ with a second bootstrap process
$X_t^{**}$ that was marginally easier to analyze.  Needless to say, this
approach caused some confusion so I removed those intermediate steps,
resulting in a clearer argument.  Other changes in the same vein have
been made as well.

I have included parts of this letter in my replies to individual
referees.  Thank you again for this opportunity.  I look forward to
your response.