### 2015-12-29 subs/ETr1/reply_editor.tex
Add response letters to editor and referees

\documentclass[12pt]{article}
\usepackage[charter]{mathdesign}
\usepackage[margin=1.25in]{geometry}
\frenchspacing
\usepackage{amsmath}

\begin{document}
\hfill 2014-10-06

\noindent
Dear Robert,

Please accept this submission of a revised version of my paper,
``Block bootstrap consistency under weak assumptions.'' I have tried to
address all of the issues raised by the referees. A short summary of
the changes to the paper follows and, as you requested, I have
written detailed letters to referees 2 and 3 addressing their specific
concerns.

\strut

\noindent \textbf{General summary of changes}

\begin{itemize}
\item Several terms were defined in the beginning of the proofs of
  Theorems~1 and~2, but were used repeatedly in the supporting
  lemmas. The paper now defines those terms in a separate subsection
  that proceeds the proofs.

\item I have made some of the arguments in the proof of Theorem~1 more
  explicit, and have tried to explain in detail how conditioning
  affects the intermediate steps of the proof.

\item One referee noticed that the proof of stochastic equicontinuity
  was in error; it has been fixed. Fixing this proof revealed an error
  in our treatment of the expectation terms, $\mu_{nt}$, and we have
  strengthened our assumptions on their dispersion around the grand
  mean, $\bar \mu_n$. We now require
  $\sum_{t=1}^n (\mu_{nt} - \bar \mu_n) \to 0$
  as $n \to \infty$.

  However, our key assumptions on the size and moment conditions that
  the stochastic array $X_{nt}$ satisfies remain unchanged, and the
  paper's main contributions are unchanged.

\item Fixing the stochastic equicontinuity proof also led me to revise
  several of the supporting Lemmas.
  \begin{itemize}
  \item Lemmas 2 and 3 of the last draft have been replaced by Lemmas
    5 and 7. These results are new maximal inequalities.
  \item Lemma 5 of the last draft has been split up; parts of it are
    in Lemma 3 and the remainder is in Lemma 4. Moreover, Lemma 4 of
    the last draft has been incorporated into Lemma 3 of the latest draft.
  \item Lemma 6 of the last draft has been rewritten to use
    unconditional expectations instead of conditional expectations.
  \end{itemize}

\item There is some concern that the inequalities referenced in the
  paper may fail with as complicated of an indexing scheme as used in
  the paper. To address this, I have added a supplemental appendix
  that provides more detail and shows how the original results apply
  to our setting. These results are presented in the main paper as
  Lemmas 6 and 7 of the paper and the supplemental appendix contains
  their proofs. The proof are essentially identical to existing
  results in the literature, and I am not claiming any authorship of
  those particular results. But some readers will probably be
  interested in this explicit demonstration.

  This seemed like the best way to address some of referee 2's
  concerns, and I have added a strong disclaimer to the appendix's
  introduction to avoid the appearance of plagarism. If you have other
  suggestions, or feel that the disclaimer is not clear enough, I am
  happy to remove the supplemental appendix and use a different
  approach.

\item Other minor changes have been made to address points raised by
  the referees.
\end{itemize}

Best,

Gray Calhoun

\end{document}