### 2015-12-31 subs/ETr2/reply_3.tex

\documentclass[12pt]{article}
\frenchspacing
\input{../../preamble.tex}
\usepackage[tiny]{titlesec}

\begin{document}
\section*{\hfill Reply to referee 3\hfill}

\begin{enumerate}
\item \textit{The first [issue] concerned the use of array notation,
    in contexts where it appeared to me unnecessary. I asked
    specifically, either for this to be dropped, or for examples to be
    given showing its role in applications.}

\item \textit{The second [issue] concerned the specification of the
    statistical model, in particular, to define the probability space
    appropriately, so as to develop the formal relationships between
    the various sub-sigma fields, and sequences and filtrations of
    same. For example, we need a formal context for objects like the
    sequence defined in equation (9). Clarity and formalism in these
    matters are especially important where random events generated by
    the bootstrap and by nature interact.}

\end{enumerate}
\end{document}