### 2015-12-31 subs/ETr2/reply_ed.tex
[fixup] Add basic response letter to editor

\documentclass[12pt]{article}
\usepackage[margin=1.25in]{geometry}
\frenchspacing
\input{../../preamble.tex}
\usepackage[tiny]{titlesec}

\begin{document}
\noindent Hi Robert,

I have revised my paper, ``Block bootstrap consistency under weak
assumptions'' (ET \#2868) as requested by the referees. The letters to
each referee addresses their requests specifically. As a quick
summary, I've made the following changes:

\input{summary}
\end{document}