### 2015-12-31 subs/ETr2/reply_2.tex

\documentclass[12pt]{article}
\frenchspacing
\input{../../preamble.tex}
\usepackage[tiny]{titlesec}

\begin{document}
\section*{\hfill Reply to referee 2\hfill}

\section*{General summary of changes}
This version of the paper has the following changes from the
previous submission:
\input{summary}

\section*{Reply to specific points}

\begin{enumerate}
\item \textit{Corollary 1: Assuming there exists an estimator
$\hat{\sigma}_n^2(\gamma)$ such that $\sup_{\gamma}
|\sigma_n^2(\gamma)/\hat{\sigma}_n^2(\gamma) - 1| \to^p 0$ may not be
innocent. After all, for an i.i.d.\ sequence of mean zero, finite
variance $\varepsilon_i$, we could have
\[
\hat{\sigma}_{n}^2(\gamma) = \tfrac{1}{n} \sum_{i=1}^{[\gamma n]} \varepsilon_i^2
\quad\text{and}\quad
\sigma_n^2(\gamma) = [\gamma n] \sigma^2 / n.
\]
Then
\begin{align*}
\sup_{0 \leq \gamma \leq 1} | \sigma_n^2(\gamma) / \hat{\sigma}_n^2(\gamma) - 1 |
& = \sup_{0 \leq \gamma \leq 1} \Big| \sigma^2 \Big/ ([n\gamma])^{-1} \sum_{i=1}^{[\gamma n]} \varepsilon_i^2 - 1 \Big| \\
& = \sup_{k:1 \leq k \leq n} \Big| \sigma^2 \Big/ k^{-1} \sum_{i=1}^{k} \varepsilon_i - 1 \Big|
\end{align*}
and the last expression converges a.s. to a non-degenerate random
variable. It is not clear to me what exactly the author wants to
achieve by dividing by $\hat{\sigma}_n(\gamma)$. Clearly such a
division achieves that the resulting process is $N(0,1)$ and not
$N(0,\gamma)$ for each $\gamma$ which means that the limit is not
Wiener measure, and therefore I did not understand how the resulting
process can end up at $W^{*}$, which is undefined, but is suggested
through the use of the letter $W$ to be some type of Wiener measure.}

\item \textit{p. 11, first equation: there is an expectation of an
    absolute value there; should the absolute value be there?}

  No, that was a mistake and it has been fixed. Thank you for pointing
  it out.

\item p. 12, Lemma 1: typo: ``$\sum_{i=1}^{J_n-1} < n$'' should read
  ``$\sum_{i=1}^{J_n-1} M_{ni} < n$''.

  Thanks!

\item \textit{On page 7, I could not verify that (18) holds; that is,
    I could not see that none of the $X_{nt}^*$ is not counted twice
    or excluded in the $\sum_{j=1}^{J_n} Z_{nj}^*$.}

  Each $Z_{nj}^*$ equals $(1/\sqrt{n}) \sum_{t=K_{n,j-1}+1}^{K_{nj}}
  (X_{nt}^* - \bar X_n)$ since $I_n(K_{n,j-1}, M_{nj}) =
  \{K_{n,j-1}+1,\dots,K_{nj}\}$. So the summations are taken over
  sequential and nonoverlapping blocks of the index $\{1,\dots,n\}$,
  starting with $\{1,\dots,M_{n1}\}$ and ending with
  $\{K_{n,J_n-1}+1,\dots,n\}$.  I've added some more details after the
  definition of $I_n(\tau,m)$ and spelled out the
  definition of $Z_{nj}^*$ a bit more to try to make
  this point more clearly in the paper. Be aware that the exact
  Equation number has moved.

\item \textit{At this time, I still do not understand the reasoning of
    the author when it comes to conditioning issues. The author to his
    credit has added explanations, and very likely the problem is
    simply my lack of experience with bootstrap issues. However, on
    page 9 for example, the author says that Lemma 5 implies that
    there exists a finite, monotone function $B(\cdot)$ that has a
    limit of 0. As I read Lemma 5, I would think that this $B(\cdot)$
    depends on $\mathcal{M}$. Similarly, on page 15, the author
    applies Lemma 6, Equation 37; yet that result was derived for an
    unconditional expectation appearing on the spot where the author
    now uses $E_{\mathcal{M}}$. In short, I am still confused by some
    of the conditioning issues, but am willing to assume that the
    problem is with me and not with the author.}

  The important point to keep in mind is that $\mathcal{M}$ only
  contains information about the block lengths for the stationary
  bootstrap. Lemmas 5 and 6 present results for the unconditional
  expectation that hold for \emph{every fixed} block length ($m'$ in
  Lemma 5, $m$ in Lemma 6). Uniform integrability over the family of
  block lengths (as in Lemma 5) implies that the same $B(\cdot)$
  applies to every $m'$ in $1,\dots,n$. Then on page 9, we use the
  fact that the same $B(\cdot)$ works for every $m'$ implies that it
  works for a randomly chosen $m'$ as well.

  I've added an intermediate step in that application of Lemma 5 to
  try to make the argument less opaque.

  The use of Lemma 6 is similar. In that expression, $M_{nj}$ is
  measurable with respect to $\mathcal{M}_n$, so it is treated the
  same as the constant $m$ in the statement of Lemma 6, and the other
  random variables are independent of $\mathcal{M}_n$. I have added
  some additional explanation that (I hope!) make this point more clear.

\end{enumerate}
\end{document}