### 2016-05-16 subs/ETr3/reply_3.tex
[#19] Add empty letter for response to referee 3
This commit adds a letter with referee 3's specific concerns, so that I
can fill out a response for the referee. This commit closes #19.

\documentclass[12pt]{article}
\frenchspacing
\input{../../preamble.tex}
\usepackage[tiny]{titlesec}

\begin{document}
\section*{\hfill Reply to referee 3\hfill}

In your letter, you wrote,
\begin{quotation}
  \noindent\textit{``Yes, your array notation is of course needed to
    accommodate the assumption of mean heterogeneity. However, what I
    was really hoping for was some additional transparency in the form
    of an example of a mean array which would satisfy Assumption 2 of
    Theorem 1, and also perhaps a case which would fail to satisfy
    it.}

  \textit{``Array notation has commonly been used in asymptotic
    analysis to increase the generality of results, especially by
    allowing flexible scaling by sample size such as
    $X_{nt} = X_t/a_n$ where $a_n$ can be either $\sqrt{n}$ or
    something else. The object here is quite different, such cases
    would not satisfy Assumption 2 and are ruled out. It would
    therefore help the general reader, who is not inside your head at
    the outset, to explain the role of the arrays in this context,
    most helpfully by quoting an example or two. Footnote 6 is not
    helpful in this respect. Lots of people like to read papers in ET
    even when they are not particularly familiar with the precedent
    literature. Giving such readers the relevant background is just a
    courtesy. We are only talking 2 or 3 extra sentences at most.''}
\end{quotation}

You also wrote,
\begin{quotation}
  \noindent\textit{``Just one additional comment on the text. On page
    5, line --7, should $\mu_n^*$ read $\bar{\mu}_n^*$?''}
\end{quotation}
\end{document}