### 2017-01-10 subs/ETr4/reply_4.tex

\documentclass[12pt]{article}
\frenchspacing
\input{../../preamble.tex}
\usepackage[tiny]{titlesec}

\begin{document}
\section*{\hfill Reply to referee 4\hfill}

\emph{On page 10, I did not understand Equation (25)\dots}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: