### 2017-01-10 subs/ETr4/reply_ed.tex

\documentclass[12pt]{article}
\frenchspacing
\input{../../preamble.tex}
\usepackage[tiny]{titlesec}

\begin{document}
\section*{\hfill Cover letter for the editor\hfill}

Hi Robert,

\strut

\noindent --- Gray Calhoun
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: