### 2017-03-27 subs/ETr4/reply_4.tex
[#38/ui] Address referee 4's comment on UI in theorem 1

\documentclass[12pt]{article}
\frenchspacing
\input{../../preamble.tex}
\usepackage[tiny]{titlesec}

\begin{document}
\section*{\hfill Reply to referee 4\hfill}

\emph{On page 10, I did not understand Equation (25)\dots}

I have rewritten this part of the proof as well as Lemma 6.  Lemma 6's
proof ignored one key measurability requirement and has been entirely
rewritten.

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: