### 2017-04-05 subs/ETr4/reply_4.tex
Write up summary of changes for resubmission

\documentclass[12pt]{article}
\frenchspacing
\input{../../preamble.tex}
\usepackage[tiny]{titlesec}

\begin{document}
\section*{\hfill Reply to referee 4\hfill}

\emph{On page 10, I did not understand Equation (25)\dots}

\strut

\noindent I have rewritten this part of the proof as well as a similar
construction in the proof of Theorem 2. Lemma 6's proof had ignored
one key measurability requirement and has been entirely rewritten as
well.

In addition, I have corrected several typos and made minor changes
suggested by the other referees.

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: