### 2017-04-05 subs/ETr4/reply_ed.tex

\documentclass[12pt]{article}
\frenchspacing
\input{../../preamble.tex}
\usepackage[tiny]{titlesec}

\begin{document}
\section*{\hfill Cover letter for the editor\hfill}

Hi Robert,

I have made the changes requested by the referees. In addition,
referee 4's concerns led me to reexamine Lemma 6 and I found a subtle
error in its proof --- the measurability requirements of some of the
variables used in the proof were inconsistent with the martingale
difference sequence arguments I was using. So I have substantially
rewritten the proof to use a different argument.

To make the review process easier, I have also included a pdf of the
paper with the changes since the last version highlighted. Thanks
again for the opportunity to revise this paper.

\strut

\noindent Best,

\strut

\noindent --- Gray Calhoun
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: